Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2020-07-10 paper/ifo-appendix.tex
tasks 1, 3, and 5

Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram. 
It is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}. A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}. This design was based off \jam{standard examples (I have no idea where I or Alex originally got the designs from, I'm confident though that they are common enough to not need citing given that we also give the diagram)} \han{[from hannah: hmm, okay - I think if we refer to `standard examples' we need to say what that means. Are there any equivalent references you can find which include the same information, even if you can't find exactly the one you used? If we think about someone who may like to reproduce this work, this is a very useful piece of information]} of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.

>>>>

Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram. 
It is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}. A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}. The design was based off standard examples found online of photo-detectors~\cite{photodiodecircuits1:online,photodiodecircuits2:online}, connecting an ADC to a Pi~\cite{piadcguide:online}, and Sallen-Key filters~\cite{sallenkey1:online,sallenkey2:online}. Useful throughout was the Pi documentation~\cite{pidocumentation:online} and the pinoutXYZ reference~\cite{pipinout:online}.

### 2020-07-10 paper/ifo-complexAudio.tex
tasks 1, 3, and 5

The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online}, which provides a convenient means to record the photodiode data. 
\han{[from hannah: James does this sound ok? And I have referenced the Raspberry Pi general website, but could you add any appropriate references you used]} \jam{[from James: no problem, sounds good. no obvious references spring to mind that aren't readily available for anyone trying to reproduce the experiment]} \han{[could you add a couple of useful resources? We should try our best to help out the reader in terms of reproducibility]}
Together, these sample the signal at roughly $16\,{\rm kHz}$.

@@

We find that the raw output of the optical microphone, collected without any further processing (except anti-aliasing), is noisy, with a loud bass hum throughout.
Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic. \jam{[non-linear appendix, Andrew says to cite Feynman lectures on non-linear oscillations]}
We find that the mains power signal is present in data recorded without an injected audio signal as seen in the power spectral density in Figure~\ref{fig:psd_noise}.
The mains power is also present in data taken with the photodiode in complete darkness with a similar but weaker distribution as that shown in Figure~\ref{fig:psd_noise}).
The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination. 
Outside the mains noise, the power appears to peak broadly around 750 Hz, the origin or significance of which we don't know.

@@

Band-passing the signal similarly does not address the remaining harmonics above $100\,{\rm Hz}$. 
However, it does suppress other noise at higher frequencies. 
The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}). 
For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\varepsilon$, the frequency response (attenuation at each frequency) as a function of frequency $\omega$ is:
\begin{equation}
\label{eq:butterworth}
H(\omega) = \left(1+\varepsilon^2 \left( \frac{\omega}{\omega_c} \right)^{2n}\right)^{-\frac{1}{2}}
\end{equation}

\jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
To construct a Butterworth band-pass filter, the low-pass filter in Eq.~\ref{eq:butterworth} is combined with a high-pass filter of similar form. 
The response of the filter used is shown in Figure~\ref{fig:butterworth}. 
The Butterworth filter performs the well as it is optimised to be `maximally flat' within the band region (prioritising condition (ii) above).
This filter improves on the above simpler filters but recordings of speech still have low intelligibility.
They can be recognised as speech without being understood.
Looking forward to Sec.~\ref{sec:logmmse}, results for the Butterworth filter are shown in combination with a speech enhancement technique.

>>>>

The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2~\cite{RaspberryPi:online}, which provides a convenient means to record the photodiode data.
Together, these sample the signal at roughly $16\,{\rm kHz}$. Resources for using the Pi are found in Appendix~\ref{app:circuit_diagram}.

@@

We find that the raw output of the optical microphone, collected without any further processing (except anti-aliasing), is noisy, with a loud bass hum throughout.
Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic. \jam{[non-linear appendix, Andrew says to cite Feynman lectures on non-linear oscillations]}
This mains power signal is present in the power spectral density of background noise as shown in the top panel of Figure~\ref{fig:psd_noise}.
The mains power signal is far weaker but still present in the background noise taken with the photodiode in darkness, suggesting that ambient lighting has a large contribution.
The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination. 
Outside the mains noise, the power appears to peak broadly around 750 Hz, the origin or significance of which we don't know.

@@

For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\varepsilon$, the frequency response (attenuation at each frequency) as a function of frequency $\omega$ is:
\begin{equation}
\label{eq:butterworth}
H(\omega) = \left(1+\varepsilon^2 \left( \frac{\omega}{\omega_c} \right)^{2n}\right)^{-\frac{1}{2}}
\end{equation}

\jam{[Andrew wanted a quantitative comment about what the Butterworth does to the PSD, what do I use?]}
To construct a Butterworth band-pass filter, the low-pass filter in Eq.~\ref{eq:butterworth} is combined with a high-pass filter of similar form. 
The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}). 
The response of the filter used is shown in Figure~\ref{fig:butterworth} and its effect on the power spectral density of background noise is shown in the bottom panel of Figure~\ref{fig:psd_noise}.
Band-passing the signal does not address the remaining harmonics above $100\,{\rm Hz}$. 
However, it does suppress other noise at higher frequencies. 
The Butterworth filter performs well at maintaining source intelligibility as it is optimised to be `maximally flat' within the band region (prioritising condition (ii) above).
However recordings of speech still have low intelligibility (the speech can't be understood) as the filter does not improve beyond the source.
% They can be recognised as speech without being understood.
Looking forward to Sec.~\ref{sec:logmmse}, results for the Butterworth filter are shown in combination with a speech enhancement technique to address this.

### 2020-07-10 paper/ifo-conclusions.tex
tasks 1, 3, and 5

In Sections~\ref{sec:ifo} and ~\ref{sec:single_tone} we demonstrate the recovery of signals with constant and slowly wandering frequency. 
A webcam sampling at $30\,{\rm Hz}$ is used to track the interference pattern and the injected signal is recovered succesfully with the Viterbi algorithm. 
These sections direcly demonstrate the continuous wave search techniques used in gravitational-wave data analysis. 
%Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer. We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.

>>>>

In Sections~\ref{sec:ifo} and ~\ref{sec:single_tone} we demonstrate the recovery of signals with constant and slowly wandering frequency. 
A webcam sampling at $30\,{\rm Hz}$ is used to track the interference pattern and the injected signal is recovered successfully with the Viterbi algorithm. 
These sections directly demonstrate the continuous wave search techniques used in gravitational-wave data analysis. 
%Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer. We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.

### 2020-07-10 paper/ifo-wanderingTone.tex
tasks 1, 3, and 5

To test the Viterbi algorithm, a signal with wandering frequency is played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}.
\jam{[The injected frequency wandered along a decaying sinusoid, it was not random.]}
%Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.

>>>>

To test the Viterbi algorithm, a sinusoidal signal with wandering frequency was played by the speaker and the output recorded via webcam as in Section~\ref{sec:single_tone}. The injected frequency itself followed a sinusoid with decaying amplitude. This choice was arbitrary, any injected path would have worked, as long as it changed slowly enough for the search parameters of the algorithm.
%Note that when creating a wandering frequency signal, the phase change must be integrated up over time, as detailed in Appendix~\ref{app:phase_gotcha}.
